A demonstration of the \LaTeXe\ class file for \itshape{SAGE Publications}

### 2015-12-22 body/overview.tex
andrea's modifications

%
\begin{figure*}
  \centering
  \begin{overpic}[width=0.8\linewidth]{figures/contact_gen}
		\put (1,1) {1} 
		\put (22,1) {2} 
		\put (42,1) {3} 
		\put (62,1) {4} 
		\put (83,1) {5} 
	\end{overpic}
  \caption{Generation of a contact configuration for the right leg of HRP-2. 1: Selection of reachable obstacles. 2: Entries of the limb samples database (with $N = 4$). 3: With a proximity query on the octree database, configurations too far from obstacles are eliminated. 4: The best candidate according to a user-defined heuristic $h$ is chosen. 5: The final contact is achieved using inverse kinematics.}
  \label{fig:contact_gen}
\end{figure*}
\subsection{Computation of a guide path --- \Pa}
We first consider the problem of planning a relevant guide path. The objective is to compute a path of root placements that will allow contact creation. A desired requirement is to preserve the completeness of the planner: it should be able to explore any possible guide trajectory, but at the same time, any computed guide trajectory must be truly feasible, i.e. must lead to a valid sequence of contacts.

\added
{
More precisely, a root placement is truly feasible if and only if there exist at least one joint configuration such that the robot satisfies three conditions: }
\begin{description}
\item[C1] is not in collision with the environment, 
\item[C2] has at least one end-effector in contact with the environment, 
\item[C3] is in static equilibrium. 
\end{description}
\added{
The conditions C1 and C2 are purely geometrical and can be efficiently approximated even without computing the contact points. Instead C3 is difficult to approximate and it requires the computation of the contact points, and the resolution of a linear program to compute the contact forces~\cite{DelPrete2016}. Moreover our intuition suggests us that, for cluttered environment, C3 is not critical, meaning that, most of the times, root placements satisfying C1 and C2 would satisfy also C3. For this reason we neglect C3 during the resolution of \Pa, and we worry about it only when solving \Pb.
}

An intuitive description of root placements satisfying C1 and C2 is ``close, but not too close'': close, because a contact surface must be partially included in the Reachable workspace of the robot (represented for the right arm in Fig.~\ref{fig:contact_gen}--1); not too close, because the robot must avoid collision (which is represented by the hull including the torso in Fig.~\ref{fig:contact_gen}--1). We define formally $C_{reach}$, the set of \replaced{root placements satisfying C1 and C2}{interesting root placements}, in which we compute a guide trajectory with a sampling-based planner, the Reachability-Based RRT---RB-RRT--- (Fig.~\ref{fig:framework}---\Pa). Planning in $C_{reach}$ boils down to planning in SE(3), which has an acceptable practical complexity.
%
Details are presented in Section~\ref{rbprm}.

@@

The second stage extends the guide path into a sequence of contact configurations (Fig.~\ref{fig:framework}--\Pb). \deleted{The true feasibility of the input path is guaranteed by the first step. This provides in turn a reasonable guarantee that a contact configuration can be obtained for any root placement along the path. }
With a fixed root position, the dimensionality of the 
contact generation problem is reduced to the number of degrees of freedom of the considered limb. In other words, we consider each limb as a manipulator attached to the root, and select the most relevant contact from a database of precomputed configurations, which is independent from the environment (Figure~\ref{fig:contact_gen}).

From a given start configuration, the planner proceeds in an iterative fashion along the discretized path: given a new root placement, an inverse-kinematics solver 
is used to maintain the contacts that were previously existing (TODO Figure). Often, these contacts cannot be maintained, because of joint limits or collisions (TODO Figure).
In such a case the contacts are broken. Conversely, new contacts are created to ensure the static equilibrium of the robot (TODO FIGURE).
The algorithm is designed so that only one contact can be created or broken between two successive configurations. This \replaced{ensures}{can be seen as a heuristic to ensure} that the interpolation between two configurations is achievable.
Details are presented in Section~\ref{sec:efort}.

To select the ``most relevant'' contact configurations, user-defined heuristics can be chosen. In Section~\ref{heuristics} we present extensively two such heuristics, which have been used in our experiments.

@@

\medskip
\textbf{A robot} is a kinematic chain $R$, comprising \mbox{$n + 6$} degrees of freedom (DOFs).
$R$ is composed of $l$ limbs $R^k, 1 \leq k \leq l$, attached to a root.
It is described by a configuration $\mathbf{q} \in SE(3) \times \mathbb{R}^n$.
We define some relevant projections of $\mathbf{q}$:
\begin{itemize}
	\item $\mathbf{q}^k$ denotes the configuration (a vector of joint values) of the limb $R^k$ (Fig.~\ref{fig:character});
	\item $\mathbf{q}^{\overline{k}}$ denotes the vector of joint values of R \textbf{not} related to $R^k$. We define for convenience \mbox{$\mathbf{q}= \mathbf{q}^k \oplus \mathbf{q}^{\overline{k}}$}; %Abusively, we denote by $\mathbf{q}^0 \oplus \mathbf{q}^k$ any configuration respecting $\mathbf{q}^0$ and $\mathbf{q}^k$;
	\item $\mathbf{q}^{0}\in SE(3)$ denotes the position and orientation of the root of the robot $R$.
\end{itemize}
%Finally, $ \mathbf{q}_{start}$ and  $\mathbf{q}_{goal}$ are the start and goal configurations given as an input of our problem.

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
The volume encompassing the trunk of the robot is denoted $W^0$ (Fig.~\ref{fig:character}-right: central cylinder). The reachable workspace of a limb $R^k$ is denoted $W^k$ (Fig.~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \, \mathbf{q}^k, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\},
\end{equation}
\adnote{in this definition you don't account for joint limits}
where $\mathbf{p}^k$ denotes the end-effector position of $R^k$ (translation only) for $\mathbf{q}^0$ being the null displacement. We also define $W = \bigcup_{k=1}^{l}W^k$, and
$W^k(\mathbf{q}^{0})$ (for $1 \leq k \leq l$) as the volume $W^k$ translated and rotated by the rigid displacement~$\mathbf{q}^{0}$.

>>>>

%
\begin{figure*}
  \centering
  \begin{overpic}[width=0.8\linewidth]{figures/contact_gen}
		\put (1,1) {a} 
		\put (22,1) {b} 
		\put (42,1) {c} 
		\put (62,1) {d} 
		\put (83,1) {e} 
	\end{overpic}
  \caption{Generation of a contact configuration for the right leg of HRP-2. a: Selection of reachable obstacles. b: Entries of the limb samples database (with $N = 4$). c: With a proximity query on the octree database, configurations too far from obstacles are eliminated. d: The best candidate according to a user-defined heuristic $h$ is chosen. e: The final contact is achieved using inverse kinematics.}
  \label{fig:contact_gen}
\end{figure*}
\subsection{Computation of a guide path --- \Pa}
We first consider the problem of planning a relevant guide path. The objective is to compute a path of root placements that will allow contact creation. A desired requirement is to preserve the completeness of the planner: it should be able to explore any possible guide trajectory, but at the same time, any computed guide trajectory must be truly feasible, i.e. must lead to a valid sequence of contacts.

More precisely, a root placement is truly feasible if and only if there exists at least one joint configuration such that the robot satisfies three conditions:
\begin{description}
\item[C1] is not in collision with the environment, 
\item[C2] has at least one end-effector in contact with the environment, 
\item[C3] is in static equilibrium. 
\end{description}
The conditions C1 and C2 are purely geometrical and can be efficiently approximated even without computing the contact points. Instead C3 is difficult to approximate and it requires the computation of the contact points, and the resolution of a linear program to compute the contact forces~\citep{Prete2016}. Moreover our intuition suggests us that, for cluttered environment, C3 is not critical, meaning that, most of the times, root placements satisfying C1 and C2 would satisfy also C3. For this reason we neglect C3 during the resolution of \Pa, and we worry about it only when solving \Pb.

An intuitive description of root placements satisfying C1 and C2 is ``close, but not too close'': close, because a contact surface must be partially included in the Reachable workspace of the robot (represented for the right arm in Figure~\ref{fig:contact_gen}--a); not too close, because the robot must avoid collision. We define formally $C_{reach}$, the set of root placements satisfying C1 and C2, in which we compute a guide trajectory with a sampling-based planner, the Reachability-Based RRT---RB-RRT--- (Figure~\ref{fig:framework}---\Pa). Planning in $C_{reach}$ boils down to planning in SE(3), which has an acceptable practical complexity.
%
Details are presented in Section~\ref{rbprm}.

@@

The second stage extends the guide path into a sequence of contact configurations (Figure~\ref{fig:framework}--\Pb). In this
step we assume that the input guide path is truly feasible.
With a fixed root position, the dimensionality of the 
contact generation problem is reduced to the number of degrees of freedom of the considered limb. In other words, we consider each limb as a manipulator attached to the root, and select the most relevant contact from a database of precomputed configurations, which is independent from the environment (Figure~\ref{fig:contact_gen}).

From a given start configuration, the planner proceeds in an iterative fashion along the discretized path: given a new root placement, an inverse-kinematics solver 
is used to maintain the contacts that were previously existing (TODO Figure). Often, these contacts cannot be maintained, because of joint limits or collisions (TODO Figure).
In such a case the contacts are broken. Conversely, new contacts are created to ensure the static equilibrium of the robot (TODO FIGURE).
The algorithm is designed so that only one contact can be created or broken between two successive configurations. This ensures that the interpolation between two configurations is achievable.
Details are presented in Section~\ref{sec:contact}.

To select the ``most relevant'' contact configurations, user-defined heuristics can be chosen. In Section~\ref{sec:heuristics} we present extensively two such heuristics, which have been used in our experiments.

@@

\medskip
\textbf{A robot} is a kinematic chain $R$, comprising \mbox{$n + 6$} degrees of freedom (DOFs).
$R$ is composed of $l$ limbs $R^k, 1 \leq k \leq l$, attached to a root.
It is described by a configuration $\mathbf{q} \in SE(3) \times \mathbb{R}^n$.
We define some relevant projections of $\mathbf{q}$:
\begin{itemize}
	\item $\mathbf{q}^k$ denotes the configuration (a vector of joint values) of the limb $R^k$ (Figure~\ref{fig:character});
	\item $\mathbf{q}^{\overline{k}}$ denotes the vector of joint values of R \textbf{not} related to $R^k$. We define for convenience \mbox{$\mathbf{q}= \mathbf{q}^k \oplus \mathbf{q}^{\overline{k}}$}; %Abusively, we denote by $\mathbf{q}^0 \oplus \mathbf{q}^k$ any configuration respecting $\mathbf{q}^0$ and $\mathbf{q}^k$;
	\item $\mathbf{q}^{0}\in SE(3)$ denotes the position and orientation of the root of the robot $R$.
\end{itemize}
%Finally, $ \mathbf{q}_{start}$ and  $\mathbf{q}_{goal}$ are the start and goal configurations given as an input of our problem.

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
The volume encompassing the trunk of the robot is denoted $W^0$ (Figure~\ref{fig:character}-right: central cylinder). The reachable workspace of a limb $R^k$ is denoted $W^k$ (Figure~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \, \mathbf{q}^k \in C^k_{jl}, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\},
\end{equation}
%~ \adnote{in this definition you don't account for joint limits}
where $\mathbf{p}^k$ denotes the end-effector position of $R^k$ (translation only) for $\mathbf{q}^0$ being the null displacement, and  $C^k_{jl}$ is the space
of admissible limb joint configurations. We also define $W = \bigcup_{k=1}^{l}W^k$, and
$W^k(\mathbf{q}^{0})$ (for $1 \leq k \leq l$) as the volume $W^k$ translated and rotated by the rigid displacement~$\mathbf{q}^{0}$.

### 2015-12-22 body/rbprm.tex
andrea's modifications

We consider the issue of computing a normalized guide path $\mathbf{q}^0(t) : [0,1] \longrightarrow SE(3)$ for the geometrical \adnote{why the root is geometrical? Isn't the path to be geometrical?} root of a multiped robot, connecting \adnote{user-defined} start and goal configurations. As said in the previous section, the goal is to enforce that any configuration $\mathbf{q}^0$ of the guide is \replaced{geometrically}{truly} feasible, i.e. can be mapped to a configuration in contact. We denote by $C_{contact} \subset SE(3) \times \mathbb{R}^n$ the contact submanifold of the robot\added{, that is the set of configurations in which the robot is in contact with the environment}.
%~ We say that a root placement $\mathbf{q}^{0}$ is \textit{truly feasible} if there exists $\mathbf{q}^{\overline{0}}$ such that 
%~ $$\mathbf{q}^{0} \oplus \mathbf{q}^{\overline{0}} \in C_{contact} $$
%~ The set of all truly-feasible root placements is denoted by $C_{reach}$. By extension, a trajectory $\mathbf{q}^0(t)$ is truly feasible if $\forall t \in [0,1], \mathbf{q}^0(t) \in C_{reach}$. 
%~ We define a bijective mapping $\pi$ between $C_{reach}$ and $C_{contact}$:
%~ \begin{equation}
    %~ \pi\colon \quad \mathbf{q}^{0} \in C_{reach} \longrightarrow  \mathbf{q}^{0} \oplus \mathbf{q}^{\overline{0}} \in C_{contact} 
    %~ \label{eq:pi}
%~ \end{equation}

We say that a root placement $\mathbf{q}^{0}$ is \textit{\replaced[remark={I don't replace 'truly feasible' with 'geometrically feasible' from now on because I don't know whether you like it}]{geometrically}{truly} feasible} if there exists at least one bijective mapping $\pi$
such that
\begin{equation}
    \pi\colon \quad \mathbf{q}^{0} \in  SE(3) \longrightarrow  \mathbf{q}^{0} \oplus \mathbf{q}^{\overline{0}} \in C_{contact} 
    \label{eq:pi}
\end{equation}

The set of all truly feasible root placements is denoted by $C_{reach}$. By extension, a path $\mathbf{q}^0(t)$ is truly feasible if $\forall t \in [0,1], \mathbf{q}^0(t) \in C_{reach}$.

For a two-stage acyclic contact planner to be exact and complete, we need the combination of two conditions on a guide trajectory generator: all the generated trajectories must be \textit{truly feasible} (sufficient condition); the guide planner must be complete, i.e. it must be able to discover any truly feasible trajectory (necessary condition).

@@

\subsection{Conditions for true feasibility}
By default, the true feasibility implies a constructive demonstration by exhibiting a valid $\pi$. This is the approach chosen by \cite{Bouyarmane2009}. However, computing a valid  $\mathbf{q}^{\overline{0}}$ is costly in practice. In this section we rather define a necessary condition and a sufficient condition for true feasibility that do not require this explicit computation.

@@

\subsubsection*{True feasibility: necessary condition}
For a contact to be possible, a volume $O_i \in O$ necessarily intersects with the reachable workspace $W(\mathbf{q}^{0})$ (Fig.~\ref{fig:contact_gen}--1). Furthermore, if $\mathbf{q}^{0}$ is truly feasible, then the trunk of the robot $W^0(\mathbf{q}^{0})$ is necessarily not colliding  with the environment $O$.

@@

\subsubsection*{\replaced[remark={for coherence with previous subsection title}]{True feasibility: sufficient condition}{Sufficient condition for true feasibility}}
%~ In building $C_{reach}^1$, we directly considered the including hull of the root body $W^{0}$, obtaining a necessary condition. 
A trivial sufficient condition for true feasibility can be constructed as a variation of $C_{reach}^1$, by replacing $W^0$ with a bounding volume $B^{max}$ encompassing the whole robot in a given pose, except for the effector surfaces to be in contact (This condition is trivial in the sense that the resulting $W$ has a zero measure. For the need of the proof, the trivial sufficient condition is enough. In practice, the construction of a non-trivial including shape $W^0$ was possible for all the robot structures we considered). \adnote{Isn't this parenthesis a bit too long?} We denote by \mbox{$C_{reach}^\infty \subset C_{reach}$} the set of root placements corresponding to the sufficient condition.

@@

Thanks to this small set of modifications, the high dimensional truly feasible path planning problem is reduced to a standard geometry and collision checking problem, in a much lower dimension (6 for Hyq, 8 for HRP-2 that has \replaced{joints in the}{a flexible} torso). \replaced{By doing do}{In so doing} we make the problem tractable for sampling-based approaches, with extremely efficient computation times.

>>>>

We consider the issue of computing a normalized guide path $\mathbf{q}^0(t) : [0,1] \longrightarrow SE(3)$ for the root of a multiped robot, connecting user-defined start and goal configurations. As said in the previous section, the goal is to enforce that any configuration $\mathbf{q}^0$ of the guide is geometrically feasible, i.e. can be mapped to a configuration in contact. We denote by $C_{contact} \subset SE(3) \times \mathbb{R}^n$ the contact submanifold of the robot, that is the set of configurations for which the robot is in contact with the environment.
%~ We say that a root placement $\mathbf{q}^{0}$ is \textit{truly feasible} if there exists $\mathbf{q}^{\overline{0}}$ such that 
%~ $$\mathbf{q}^{0} \oplus \mathbf{q}^{\overline{0}} \in C_{contact} $$
%~ The set of all truly-feasible root placements is denoted by $C_{reach}$. By extension, a trajectory $\mathbf{q}^0(t)$ is truly feasible if $\forall t \in [0,1], \mathbf{q}^0(t) \in C_{reach}$. 
%~ We define a bijective mapping $\pi$ between $C_{reach}$ and $C_{contact}$:
%~ \begin{equation}
    %~ \pi\colon \quad \mathbf{q}^{0} \in C_{reach} \longrightarrow  \mathbf{q}^{0} \oplus \mathbf{q}^{\overline{0}} \in C_{contact} 
    %~ \label{eq:pi}
%~ \end{equation}

We say that a root placement $\mathbf{q}^{0}$ is geometrically feasible if there exists at least one bijective mapping $\pi$
such that
\begin{equation}
    \pi\colon \quad \mathbf{q}^{0} \in  SE(3) \longrightarrow  \mathbf{q}^{0} \oplus \mathbf{q}^{\overline{0}} \in C_{contact} 
    \label{eq:pi}
\end{equation}

The set of all geometrically feasible root placements is denoted by $C_{reach}$. By extension, a path $\mathbf{q}^0(t)$ is truly feasible if $\forall t \in [0,1], \mathbf{q}^0(t) \in C_{reach}$.

For a two-stage acyclic contact planner to be exact and complete, we need the combination of two conditions on a guide trajectory generator: all the generated trajectories must be \textit{geometrically feasible} (sufficient condition); the guide planner must be complete, i.e. it must be able to discover any truly feasible trajectory (necessary condition).

@@

\subsection{Conditions for geometric feasibility}
By default, the geometric feasibility implies a constructive demonstration by exhibiting a valid $\pi$. This is the approach chosen by \cite{Bouyarmane2009}. However, computing a valid  $\mathbf{q}^{\overline{0}}$ is costly in practice. In this section we rather define a necessary condition and a sufficient condition for geometric feasibility that do not require this explicit computation.

@@

\subsubsection*{Geometric feasibility: necessary condition}
For a contact to be possible, a volume $O_i \in O$ necessarily intersects with the reachable workspace $W(\mathbf{q}^{0})$ (Figure~\ref{fig:contact_gen}--1). Furthermore, if $\mathbf{q}^{0}$ is  feasible, then the trunk of the robot $W^0(\mathbf{q}^{0})$ is necessarily not colliding  with the environment $O$.

@@

\subsubsection*{Geometric feasibility: sufficient condition}
%~ In building $C_{reach}^1$, we directly considered the including hull of the root body $W^{0}$, obtaining a necessary condition. 
A trivial sufficient condition for geometric feasibility can be constructed as a variation of $C_{reach}^1$, by replacing $W^0$ with a bounding volume $B^{max}$ encompassing the whole robot in a given pose, except for the effector surfaces to be in contact. We denote by \mbox{$C_{reach}^\infty \subset C_{reach}$} the set of root placements corresponding to the sufficient condition.

@@

Thanks to this small set of modifications, the high dimensional truly feasible path planning problem is reduced to a standard geometry and collision checking problem, in a much lower dimension (6 for Hyq, 8 for HRP-2 that has joints in the torso). In so doing we make the problem tractable for sampling-based approaches, with extremely efficient computation times.
